---
title: "P50_LME_visualization"
author: "Xue Zhang"
date: "7/12/2022"
output:
  html_document: default
  pdf_document: default
  word_document: default
always_allow_html: yes
---


### load envionment and packages and subfunctions
```{r setup, include=FALSE}
# use here from the here package
here <- here::here

# set up environment and load useful packages
source(here('R', 'set_up_env.R'))

# load functions for data manipulations
source(here('R', 'data_manipulation.R'))

# load functions for data visualization, including tables and figures
source(here('R', 'visualization.R'))

# load functions for folder manipulation
source(here('R','folder_manipulation.R'))
```


### define data dir and create folders for output figures
```{r define_folders}
# data folder
csvdir = "/Users/xuezhang/Documents/Stanford/Projects/P50/Analysis/CSV"

# define and create figure folder
Figuredir = here("Figures")
Figsubdir = c("TIFF","PDF")
Figsubsubdir = c("Boxplot", "OverallCorrelation", "RepeatedCorrelation")
for (subdir in Figsubdir) {
  for (subsubdir in Figsubsubdir) {
    des_dir <- paste0(Figuredir, "/", subdir, "/", subsubdir)
    create_folder_subfolder(des_dir)
  }
}

```


```{r define_parameters}
# subjects
subjn = 14 # sub num + 1
subj_list = as.character(1:(subjn-1))
subj_label = c(1:(subjn-1))

# dosage information
dosage_level <- c("Placebo", "0.05mg/kg", "0.5mg/kg")
dosage_level_2ses <- c("Placebo", "0.5mg/kg")
Dosage <- import(paste(csvdir,"/Ket_Dose_unblinded_demo.xlsx",sep = ""))

# for plot
xlabels_2time = c("Pre", "Post")

# figure related parameters
mycolor <- colorRampPalette(brewer.pal(name="Set1", n = 9))(length(subj_label))
dosecolor <- c('#B5B5B5', '#FFAD65', '#C53600')

xtitle_size = 4
title_size = 7
legend_title_size = 9
xlabel_size = 7

```


# read questionnaire data (CADSS and 5D-ASC) and FEET activation data from amygdala and anterior insula
```{r read data}
source(here('R', 'read_redcap_data.R'))
source(here('R', 'read_fmri_data.R'))
```


### LME for CADSS: y ~ Time*Dosage + Age + Gender + (1|Subjects) 
```{r fig.width=8,fig.height=4, fig.align='center',comment= NA, message=FALSE, tidy = TRUE, results='asis',warning=FALSE}
desired_columns = c("Subjects", "Time", "CADSS", "CADSS_depersonalization", "CADSS_derealization", "CADSS_amnesia", "Dosage", "Age", "Gender")

data_fortest <- merge(data_repeated_pervisit, age_gender, by = c("Subjects"), all=TRUE)

data_fortest <- data_fortest %>%
  select(desired_columns) %>%
  pivot_longer(-c(Subjects, Time, Dosage, Age, Gender), names_to = "variables", values_to = "value") %>%
  mutate(variables = factor(variables, levels = c("CADSS", "CADSS_depersonalization", "CADSS_derealization", "CADSS_amnesia")))

formula <-  as.formula(paste("value ~ Time*Dosage + Age + Gender + (1|Subjects)", sep = ""))

data_fortest %>%
split(.$variables) %>%
  map(~anova(lmer(formula,data = ., REML = FALSE), type = 'marginal')) %>%
  tibble(
    dvsub = names(.),
    untidied = .
    ) %>%
  mutate(tidy = map(untidied, broom::tidy)) %>%
  unnest(tidy) %>%
  select(-untidied) %>%
  adjust_pvalue(method = "fdr")%>%
  add_significance() %>%
  select(dvsub, term, NumDF,DenDF, statistic, p.value, p.value.adj, p.value.adj.signif) %>%
  kbl(caption = "ANOVA Marginal") %>%
  kable_classic(full_width = F, html_font = "Arial", font_size = 10)

```


### LME for 5D-ASC: y ~ Dosage + Age + Gender + (1|Subjects) 
```{r fig.width=2,fig.height=2, fig.align='center',comment= NA, message=FALSE, tidy = TRUE, results='asis',warning=FALSE}
data_fortest <- data_once_pervisit %>%
  select(Subjects, Dosage, dascscore_anxiety, dascscore_bliss, dascscore_impair) %>%
  pivot_longer(-c(Subjects, Dosage), names_to = "variables", values_to = "value") %>%
      mutate(Dosage = factor(Dosage, levels = dosage_level))

stat.test <- data_fortest %>%
  group_by(variables) %>%
  t_test(value ~ Dosage, paired = TRUE) %>% #
  add_significance(.,"p","p.signif")%>%
  add_xy_position(x = "Dosage", scales = "free") 

t_test_table <- stat.test %>%
  select(colnames(.)[c(-2, -5:-6, -10:-11, -13:-16)]) %>%
  mutate(group = paste0(group1,' vs ',  group2, sep= "")) %>%
  select(-group1, -group2) %>%
  pivot_wider(names_from = group, values_from = c(statistic, df, p)) %>%
  kbl(caption = "post-hoc t test") %>%
  kable_classic(full_width = F, html_font = "Arial", font_size = 10)


# Generate statistics

data_fortest <- merge(data_once_pervisit, age_gender, by = c("Subjects"), all=TRUE)
data_fortest <- data_fortest %>%
  select(Subjects, Dosage, Age, Gender, matches("dasc")) %>%
  pivot_longer(-c(Subjects, Dosage, Age, Gender), names_to = "variables", values_to = "value")

formula <-  as.formula(paste("value ~ Dosage + Gender + Age + (1|Subjects)", sep = ""))

data_fortest %>%
split(.$variables) %>%
  map(~anova(lmer(formula,data = ., REML = FALSE)), type = 'marginal') %>%
  tibble(
    dvsub = names(.),
    untidied = .
    ) %>%
  mutate(tidy = map(untidied, broom::tidy)) %>%
  unnest(tidy) %>%
  select(-untidied) %>%
  adjust_pvalue(method = "fdr")%>%
  add_significance(.,"p.value.adj","p.value.adj.signif") %>%
  select(dvsub, term, NumDF,DenDF, statistic, p.value, p.value.adj, p.value.adj.signif) %>%
  kbl(caption = "ANOVA Marginal") %>%
  kable_classic(full_width = F, html_font = "Arial", font_size = 10)
```

# create column plot for CADSS, 5D-ASC and FEET activations in paper
```{r fig.width=2,fig.height=2, fig.align='center',comment= NA, message=FALSE, tidy = TRUE, results='asis',warning=FALSE}
# combine all CADSS and 5D-ASC subcomponents of all three visits
non_brain_data_post_pre <- Reduce(function(x, y) merge(x, y, by = c("Subjects", "Dosage"), all=TRUE), list(data_repeated_post_pre, data_once_pervisit))

data_fortest <- non_brain_data_post_pre %>%
  select(Subjects, Dosage, CADSS_depersonalization_post_pre, CADSS_derealization_post_pre, CADSS_amnesia_post_pre, dascscore_anxiety, dascscore_bliss,dascscore_impair) %>%
  pivot_longer(-c(Subjects, Dosage), names_to = "variables", values_to = "value") %>%
      mutate(Dosage = factor(Dosage, levels = dosage_level))

# by subject by variable
plot_variables <- unique(data_fortest$variables)

plot_titles <- unique(data_fortest$variables)
figure_width = 0.7
figure_height = 1.2
boxplot_width = 0.4/figure_width
colors_network =  c("#E5BBD8", "#E5BBD8", "#ffc100", "#E5BBD8", "#ffc100", "#ffc100", "#ffc100")
xlabel_text = c("Placebo","0.05mg/kg","0.5mg/kg")
# xlabel_text = c("Placebo","0.5mg/kg")
for (ivar in 1:length(plot_variables)) {
  # generate paired t-test significance between any two of the three drug conditions
  stat.test <- data_fortest %>%
  filter(variables == plot_variables[ivar]) %>%
  group_by(variables) %>%
  t_test(value ~ Dosage, paired = TRUE) %>%
  add_significance(.,"p","p.signif")%>%
  add_xy_position(x = "Dosage", scales = "free")%>%
  filter(p < 0.05)
  
  # summarize mean and sd
  data_fortest_sta <- data_fortest %>%
    filter(variables == plot_variables[ivar]) %>%
    group_by(Dosage) %>%
    summarise(
      count = sum(!is.na(value)),
      sd = sd(value, na.rm = TRUE),
      se = sd / sqrt(count),
      avg = mean(value, na.rm = TRUE))
  
  ylim_max <- max(data_fortest_sta$avg) + 1.2 * max(data_fortest_sta$se)
  ylim_min <- min(data_fortest_sta$avg) - 1.2 * max(data_fortest_sta$se)
  
  if (ylim_min < 0 & min(data_fortest_sta$avg)>= 0) {
    ylim_min = 0
    
  }

  colors_network_plot <- c(colors_network[ivar], colors_network[ivar], colors_network[ivar])
  
  
  
      myplot <- data_fortest_sta %>%
        ggplot(aes(x = Dosage, y = avg)) +
        geom_col(aes(group = Dosage, fill = Dosage), alpha = 1, size = 0.25) +
        geom_errorbar(aes(group = Dosage, color = Dosage, ymin = avg-se, ymax = avg+se), width = 0, size = 0.25) + scale_y_continuous(name="", limits=c(ylim_min, ylim_max), breaks = seq(round(ylim_min) ,round(ylim_max), by = floor((ylim_max - ylim_min)/4)), expand = c(0, 0)) + scale_x_discrete(breaks = waiver(), labels = xlabel_text) + theme_classic() + theme(axis.ticks = element_line(size = 0.2), axis.line = element_line(size = 0.2),
    axis.text.x = element_text(size = 9), 
        axis.title.y=element_blank(),
        strip.text.x = element_text(size = 8, colour = "black"),
        panel.margin = unit(0.25, "lines"), legend.position="bottom", legend.margin=margin(t=-10),
        legend.spacing.y = unit(-0.05, 'cm')) + scale_colour_manual(values = colors_network_plot) + scale_fill_manual(values = colors_network_plot)+ guides(color = guide_legend(override.aes= list(alpha = 0.5))) + guides(col = guide_legend(nrow = 1))
      
        
      
      if (ylim_min <= 0) {
        myplot <- myplot + geom_hline(yintercept = 0, size = 0.1)
      }
  
  figure_width = 1.3
  xtitle_size = 6
  title_size = 6
  legend_title_size = 3
  myplot <- add_theme(myplot, xtitle_size, title_size, legend_title_size)
  myplot <- myplot+scale_colour_manual(values = colors_network_plot) + scale_fill_manual(values = colors_network_plot)
  leg <- get_legend(myplot)
plot_leg <- as_ggplot(leg)
  myplot <- myplot + theme(legend.position = "off")

  # TIFF
  ggsave(file=paste(Figuredir, "/TIFF/Boxplot/", plot_variables[ivar],"_col.TIFF", sep = ""), plot=myplot, width=figure_width, height=1.2, dpi = 300)
  # PDF
  ggsave(file=paste(Figuredir, "/PDF/Boxplot/", plot_variables[ivar],"_col.PDF", sep = ""),  device = cairo_pdf, plot=myplot, width=figure_width, height=1.2, dpi = 300)
}

```

### LME for FEET task activations: y ~ Dosage + Age + Gender + (1|Subjects)
### and column plot visualization
```{r fig.width=8,fig.height=6, fig.align='center',comment= NA, message=FALSE, tidy = TRUE, results='asis',warning=FALSE}

task_name <- unique(Faces_data$Task)
con_name <- unique(Faces_data$Contrast)

for (i_task in task_name) {
  for (i_con in con_name) {
    print(paste0("Task: ", i_task,"; Contrast: ", i_con,"\n"))
    
    data_fortest <- Faces_data %>%
      filter(Task == i_task, Contrast == i_con) %>%
      pivot_longer(-c(Subjects, Dosage, Task, Contrast), names_to = "variables", values_to = "value")
    
    stat.test <- data_fortest %>%
      group_by(variables) %>%
      t_test(value ~ Dosage, paired = TRUE) %>% # 
      add_significance(.,"p","p.signif")
    
    t_test_table <- stat.test %>%
  select(colnames(.)[c(-2, -5:-6, -10:-11, -13:-16)]) %>%
  mutate(group = paste0(group1,' vs ',  group2, sep= "")) %>%
  select(-group1, -group2) %>%
  pivot_wider(names_from = group, values_from = c(statistic, df, p, p.signif)) %>%
  kbl(caption = "post-hoc t test") %>%
  kable_classic(full_width = F, html_font = "Arial", font_size = 10)
  
    # run LME and generate statistics
    rois = unique(data_fortest$variables)
    formula <-  as.formula(paste("value ~ Dosage + (1|Subjects)", sep = ""))
    print(data_fortest %>%
            filter(variables %in% rois) %>% 
            split(.$variables) %>%
            map(~anova(lmer(formula,data = ., REML = FALSE)), type = 'marginal') %>%
            tibble(
              dvsub = names(.),
              untidied = .
            ) %>%
            mutate(tidy = map(untidied, broom::tidy)) %>%
            unnest(tidy) %>%
            select(-untidied) %>%
            adjust_pvalue(method = "fdr")%>%
            add_significance() %>%
            select(dvsub, NumDF,DenDF, statistic, p.value, p.value.adj) %>%
            kbl(caption = "ANOVA Marginal") %>%
            kable_classic(full_width = F, html_font = "Arial", font_size = 10))
    
   
   # generate column plots 
    plot_variables <- rois
    plot_titles <- c("antInsula R", "Amygdala R")
    figure_width = 0.7
    figure_height = 1.2
    boxplot_width = 0.4/figure_width
    xlabel_text = c("Placebo","0.05mg/kg","0.5mg/kg")
    colors_network =  c("#46B674", "#D77842")
    
    
    for (ivar in 1:length(plot_variables)) {
      
      # generate significance
      stat.test <- data_fortest %>%
        filter(variables == plot_variables[ivar]) %>%
        group_by(variables) %>%
        t_test(value ~ Dosage, paired = TRUE) %>%
        add_significance(.,"p","p.signif")%>%
        add_xy_position(x = "Dosage", scales = "free")%>%
        filter(p < 0.05)
      
      subjN = length(unique(data_fortest$Subjects))
      
      # summarize mean and sd
      data_fortest_sta <- data_fortest %>%
        filter(variables == plot_variables[ivar]) %>%
        group_by(Dosage) %>%
        summarise(
          count = sum(!is.na(value)),
          sd = sd(value, na.rm = TRUE),
          se = sd / sqrt(count),
          avg = mean(value, na.rm = TRUE))
      
      
      
      
      ylim_max <- max(data_fortest_sta$avg) + 1.1 * max(data_fortest_sta$se)
  ylim_min <- min(data_fortest_sta$avg) - 1.1 * max(data_fortest_sta$se)
      
  colors_network_plot <- c(colors_network[ivar], colors_network[ivar], colors_network[ivar])
  
     myplot <- data_fortest_sta %>%
        ggplot(aes(x = Dosage, y = avg)) +
        geom_col(aes(group = Dosage, fill = Dosage), alpha = 1, size = 0.25) +
        geom_errorbar(aes(group = Dosage, color = Dosage, ymin = avg-se, ymax = avg+se), width = 0, size = 0.25) + guides(color = guide_legend(override.aes= list(alpha = 0.5))) + scale_x_discrete(breaks = waiver(), labels = xlabel_text) + scale_y_continuous(name="Size of effect", limits=c(-0.3, 0.75), breaks = c(-0.25, 0, 0.25, 0.5, 0.75), expand = c(0, 0)) + theme_classic() + theme(axis.ticks = element_line(size = 0.2), axis.line = element_line(size = 0.2),
    axis.text.x = element_text(size = 9), 
        axis.title.y=element_blank(),
        strip.text.x = element_text(size = 8, colour = "black"),
        panel.margin = unit(0.25, "lines"), legend.position="bottom", legend.margin=margin(t=-10),
        legend.spacing.y = unit(-0.05, 'cm')) + scale_colour_manual(values = colors_network_plot) + scale_fill_manual(values = colors_network_plot) + guides(color = guide_legend(override.aes= list(alpha = 0.5))) + guides(col = guide_legend(nrow = 1)) 
      if (ylim_min < 0) {
      myplot <- myplot + geom_hline(yintercept = 0, size = 0.1)
      }
      
      xtitle_size = 6
      title_size = 6
      legend_title_size = 3
      myplot <- add_theme(myplot, xtitle_size, title_size, legend_title_size)
      myplot <- myplot+scale_colour_manual(values = colors_network_plot) + scale_fill_manual(values = colors_network_plot)
      leg <- get_legend(myplot)
      plot_leg <- as_ggplot(leg)
      myplot <- myplot + theme(legend.position = "off")
      # TIFF
      figure_width = 1.3
      ggsave(file=paste(Figuredir, "/TIFF/Boxplot/",i_task, "_",i_con, "_", plot_variables[ivar],"_col_trial.TIFF", sep = ""), plot=myplot, width=figure_width, height=1.4, dpi = 1000)
      # ggsave(file=paste(Figuredir, "/PDF/Boxplot/", i_task, "_",i_con, "_", plot_variables[ivar],"_legend_col.pdf", sep = ""), device = cairo_pdf, plot=plot_leg, width=figure_width*2, height=1.2, dpi = 300)
      # PDF
      ggsave(file=paste(Figuredir, "/PDF/Boxplot/", i_task, "_",i_con, "_", plot_variables[ivar],"_col_trial.PDF", sep = ""),  device = cairo_pdf, plot=myplot, width=figure_width, height=1.2, dpi = 1000)
 
    }

    
  }
  
}

```

