---
title: "P50_LME_visualization"
author: "Xue Zhang"
date: "7/12/2022"
output:
  html_document: default
  pdf_document: default
  word_document: default
always_allow_html: yes
---


### load envionment and packages and subfunctions
```{r setup, include=FALSE}
# use here from the here package
here <- here::here

# set up environment and load useful packages
source(here('R', 'set_up_env.R'))

# load functions for data manipulations
source(here('R', 'data_manipulation.R'))

# load functions for data visualization, including tables and figures
source(here('R', 'visualization.R'))

```

```{r define_parameters, warning=FALSE}
# data folder
csvdir = csvdir = here('csv')
# subjects
subjn = 14 # sub num + 1
subj_list = as.character(1:(subjn-1))
subj_label = c(1:(subjn-1))

# dosage information
dosage_level <- c("Placebo", "0.05mg/kg", "0.5mg/kg")
dosage_level_2ses <- c("Placebo", "0.5mg/kg")
Dosage <- import(paste(csvdir,"/Ket_Dose_unblinded_demo.xlsx",sep = ""))

# for plot
xlabels_2time = c("Pre", "Post")

# figure related parameters
mycolor <- colorRampPalette(brewer.pal(name="Set1", n = 9))(length(subj_label))
dosecolor <- c('#B5B5B5', '#FFAD65', '#C53600')

xtitle_size = 4
title_size = 7
legend_title_size = 9
xlabel_size = 7

```


# read questionnaire data (CADSS and 5D-ASC) and FEET activation data from amygdala and anterior insula
```{r read data, warning=FALSE, message = FALSE}
source(here('R', 'read_redcap_data.R'))
source(here('R', 'read_fmri_data.R'))
```


### LME for CADSS: y ~ Time*Dosage + Age + Sex + (1|Subjects) 
```{r fig.width=8,fig.height=4, fig.align='center',comment= NA, message=FALSE, tidy = TRUE, results='asis',warning=FALSE}
desired_columns = c("Subjects", "Time", "CADSS", "CADSS_depersonalization", "CADSS_derealization", "CADSS_amnesia", "Dosage", "Age", "Sex")

data_fortest <- merge(data_repeated_pervisit, age_Sex, by = c("Subjects"), all=TRUE)

data_fortest <- data_fortest %>%
  select(desired_columns) %>%
  pivot_longer(-c(Subjects, Time, Dosage, Age, Sex), names_to = "variables", values_to = "value") %>%
  mutate(variables = factor(variables, levels = c("CADSS", "CADSS_depersonalization", "CADSS_derealization", "CADSS_amnesia")))

formula <-  as.formula(paste("value ~ Time*Dosage + Age + Sex + (1|Subjects)", sep = ""))

data_fortest %>%
split(.$variables) %>%
  map(~anova(lmer(formula,data = ., REML = FALSE), type = 'marginal')) %>%
  tibble(
    dvsub = names(.),
    untidied = .
    ) %>%
  mutate(tidy = map(untidied, broom::tidy)) %>%
  unnest(tidy) %>%
  select(-untidied) %>%
  adjust_pvalue(method = "fdr")%>%
  add_significance() %>%
  select(dvsub, term, NumDF,DenDF, statistic, p.value, p.value.adj, p.value.adj.signif) %>%
  rename("y" = "dvsub", "x" = "term", "Fstats" = "statistic") %>%
  kbl(caption = "ANOVA results for CADSS LME: y ~ Time*Dosage + Age + Sex + (1|Subjects)") %>%
  kable_classic(full_width = F, html_font = "Arial", font_size = 10)

# generate paired t-test significance between any two of the three drug conditions
data_fortest <- data_repeated_post_pre %>%
  pivot_longer(-c(Subjects, Dosage), names_to = "variables", values_to = "value") 

stat.test <- data_fortest %>%
  group_by(variables) %>%
  t_test(value ~ Dosage, paired = TRUE) %>%
  add_significance(.,"p","p.signif")%>%
  add_xy_position(x = "Dosage", scales = "free") 

stat.test %>%
  select(colnames(.)[c(-2, -5:-6, -10:-11, -13:-16)]) %>%
  mutate(group = paste0(group1,' vs ',  group2, sep= "")) %>%
  select(-group1, -group2) %>%
  pivot_wider(names_from = group, values_from = c(statistic, df, p)) %>%
  kbl(caption = "suppl. table 1: post-hoc t test") %>%
  kable_classic(full_width = F, html_font = "Arial", font_size = 10)

```


### LME for 5D-ASC: y ~ Dosage + Age + Sex + (1|Subjects) 
```{r fig.width=2,fig.height=2, fig.align='center',comment= NA, message=FALSE, tidy = TRUE, results='asis',warning=FALSE}
data_fortest <- data_once_pervisit %>%
  select(Subjects, Dosage, dascscore_anxiety, dascscore_bliss, dascscore_impair) %>%
  pivot_longer(-c(Subjects, Dosage), names_to = "variables", values_to = "value") %>%
      mutate(Dosage = factor(Dosage, levels = dosage_level))

# run LME and generate statistics

data_fortest <- merge(data_once_pervisit, age_Sex, by = c("Subjects"), all=TRUE)
data_fortest <- data_fortest %>%
  select(Subjects, Dosage, Age, Sex, matches("dasc")) %>%
  pivot_longer(-c(Subjects, Dosage, Age, Sex), names_to = "variables", values_to = "value")

formula <-  as.formula(paste("value ~ Dosage + Sex + Age + (1|Subjects)", sep = ""))

data_fortest %>%
split(.$variables) %>%
  map(~anova(lmer(formula,data = ., REML = FALSE)), type = 'marginal') %>%
  tibble(
    dvsub = names(.),
    untidied = .
    ) %>%
  mutate(tidy = map(untidied, broom::tidy)) %>%
  unnest(tidy) %>%
  select(-untidied) %>%
  adjust_pvalue(method = "fdr")%>%
  add_significance(.,"p.value.adj","p.value.adj.signif") %>%
  select(dvsub, term, NumDF,DenDF, statistic, p.value, p.value.adj, p.value.adj.signif) %>%
  rename("y" = "dvsub", "x" = "term", "Fstats" = "statistic") %>%
  kbl(caption = "ANOVA results for 5D-ASC LME: y ~ Dosage + Age + Sex + (1|Subjects)") %>%
  kable_classic(full_width = F, html_font = "Arial", font_size = 10)

# generate paired t-test significance between any two of the three drug conditions
stat.test <- data_fortest %>%
  group_by(variables) %>%
  t_test(value ~ Dosage, paired = TRUE) %>% #
  add_significance(.,"p","p.signif")%>%
  add_xy_position(x = "Dosage", scales = "free") 

stat.test %>%
  select(colnames(.)[c(-2, -5:-6, -10:-11, -13:-16)]) %>%
  mutate(group = paste0(group1,' vs ',  group2, sep= "")) %>%
  select(-group1, -group2) %>%
  pivot_wider(names_from = group, values_from = c(statistic, df, p)) %>%
  kbl(caption = "suppl. table 1: post-hoc t test") %>%
  kable_classic(full_width = F, html_font = "Arial", font_size = 10)
```

# create column plot for CADSS and 5D-ASC in paper - figures in fig. 1a, 2a, 3a and suppl. fig. 1d-i
```{r fig.width=2,fig.height=2, fig.align='center',comment= NA, message=FALSE, tidy = TRUE, results='asis',warning=FALSE}
# combine all CADSS and 5D-ASC subcomponents of all three visits
non_brain_data_post_pre <- Reduce(function(x, y) merge(x, y, by = c("Subjects", "Dosage"), all=TRUE), list(data_repeated_post_pre, data_once_pervisit))

data_fortest <- non_brain_data_post_pre %>%
  select(Subjects, Dosage, CADSS_depersonalization_post_pre, CADSS_derealization_post_pre, CADSS_amnesia_post_pre, dascscore_anxiety, dascscore_bliss,dascscore_impair) %>%
  pivot_longer(-c(Subjects, Dosage), names_to = "variables", values_to = "value") %>%
      mutate(Dosage = factor(Dosage, levels = dosage_level))

# by subject by variable
plot_variables <- unique(data_fortest$variables)

plot_titles <- unique(data_fortest$variables)
figure_width = 0.7
figure_height = 1.2
boxplot_width = 0.4/figure_width
colors_network =  c("#E5BBD8", "#E5BBD8", "#ffc100", "#E5BBD8", "#ffc100", "#ffc100", "#ffc100")
xlabel_text = c("Placebo","0.05mg/kg","0.5mg/kg")
# xlabel_text = c("Placebo","0.5mg/kg")
for (ivar in 1:length(plot_variables)) {
  # summarize mean and sd
  data_fortest_sta <- data_fortest %>%
    filter(variables == plot_variables[ivar]) %>%
    group_by(Dosage) %>%
    summarise(
      count = sum(!is.na(value)),
      sd = sd(value, na.rm = TRUE),
      se = sd / sqrt(count),
      avg = mean(value, na.rm = TRUE))
  
  ylim_max <- max(data_fortest_sta$avg) + 1.2 * max(data_fortest_sta$se)
  ylim_min <- min(data_fortest_sta$avg) - 1.2 * max(data_fortest_sta$se)
  
  if (ylim_min < 0 & min(data_fortest_sta$avg)>= 0) {
    ylim_min = 0
    
  }

  colors_network_plot <- c(colors_network[ivar],
                           colors_network[ivar],
                           colors_network[ivar])
  myplot <- data_fortest_sta %>%
        ggplot(aes(x = Dosage, y = avg)) +
        geom_col(aes(group = Dosage, fill = Dosage), alpha = 1, size = 0.25) +
        geom_errorbar(aes(group = Dosage, color = Dosage, ymin = avg-se, ymax = avg+se), width = 0, size = 0.25) +  scale_x_discrete(breaks = waiver(), labels = xlabel_text) + theme_classic() + theme(axis.ticks = element_line(size = 0.2), axis.line = element_line(size = 0.2),
    axis.text.x = element_text(size = 9), 
        axis.title.y=element_blank(),
        strip.text.x = element_text(size = 8, colour = "black"),
        panel.margin = unit(0.25, "lines"), legend.position="bottom", legend.margin=margin(t=-10),
        legend.spacing.y = unit(-0.05, 'cm')) + scale_colour_manual(values = colors_network_plot) + scale_fill_manual(values = colors_network_plot)+ guides(color = guide_legend(override.aes= list(alpha = 0.5))) + guides(col = guide_legend(nrow = 1)) + ggtitle(plot_variables[ivar])
      
        
      
      if (ylim_min <= 0) {
        myplot <- myplot + geom_hline(yintercept = 0, size = 0.1)
      }
  
  figure_width = 1.3
  xtitle_size = 6
  title_size = 6
  legend_title_size = 3
  myplot <- add_theme(myplot, xtitle_size, title_size, legend_title_size)
  myplot <- myplot+scale_colour_manual(values = colors_network_plot) + scale_fill_manual(values = colors_network_plot)
  leg <- get_legend(myplot)
plot_leg <- as_ggplot(leg)
  myplot <- myplot + theme(legend.position = "off")
 print(myplot)
}

```

### LME for FEET task threat vs neutral activations: y ~ Dosage + Age + Sex + (1|Subjects) and column plot visualization (suppl. fig 1j and 1k)
```{r fig.width=2,fig.height=2, fig.align='center',comment= NA, message=FALSE, tidy = TRUE, results='asis',warning=FALSE}

task_name <- c('nonconscious')
con_name <- c('ThreatvsNeutral')

for (i_task in task_name) {
  for (i_con in con_name) {
    print(paste0("Task: ", i_task,"; Contrast: ", i_con,"\n"))
    
    data_fortest <- Faces_data %>%
      filter(Task == i_task, Contrast == i_con) %>%
      pivot_longer(-c(Subjects, Dosage, Task, Contrast), names_to = "variables", values_to = "value")

    # run LME and generate statistics
    rois = unique(data_fortest$variables)
    formula <-  as.formula(paste("value ~ Dosage  + Age + Sex + (1|Subjects)", sep = ""))
    print(data_fortest %>%
            filter(variables %in% rois) %>% 
            split(.$variables) %>%
            map(~anova(lmer(formula,data = ., REML = FALSE)), type = 'marginal') %>%
            tibble(
              dvsub = names(.),
              untidied = .
            ) %>%
            mutate(tidy = map(untidied, broom::tidy)) %>%
            unnest(tidy) %>%
            select(-untidied) %>%
            adjust_pvalue(method = "fdr")%>%
            add_significance() %>%
            select(dvsub, term, NumDF,DenDF, statistic, p.value, p.value.adj) %>%
            rename("y" = "dvsub", "x" = "term", "Fstats" = "statistic") %>%
            kbl(caption = "ANOVA results for FEET task activations LME: y ~ Dosage + Age + Sex + (1|Subjects)") %>%
            kable_classic(full_width = F, html_font = "Arial", font_size = 10))
    # running post-hoc paired t-test for suppl. table 1
  stat.test <- data_fortest %>%
      group_by(variables) %>%
      t_test(value ~ Dosage, paired = TRUE) %>% # 
      add_significance(.,"p","p.signif")
  
  stat.test %>%
  select(colnames(.)[c(-2, -5:-6, -10:-11, -13:-16)]) %>%
  mutate(group = paste0(group1,' vs ',  group2, sep= "")) %>%
  select(-group1, -group2) %>%
  pivot_wider(names_from = group, values_from = c(statistic, df, p, p.signif)) %>%
  kbl(caption = "suppl. table1: post-hoc t test") %>%
  kable_classic(full_width = F, html_font = "Arial", font_size = 10)
   
   # generate column plots 
    plot_variables <- rois
    plot_titles <- c("antInsula R", "Amygdala R")
    figure_width = 0.7
    figure_height = 1.2
    boxplot_width = 0.4/figure_width
    xlabel_text = c("Placebo","0.05mg/kg","0.5mg/kg")
    colors_network =  c("#46B674", "#D77842")
    
    
    for (ivar in 1:length(plot_variables)) {
      
      # generate significance
      stat.test <- data_fortest %>%
        filter(variables == plot_variables[ivar]) %>%
        group_by(variables) %>%
        t_test(value ~ Dosage, paired = TRUE) %>%
        add_significance(.,"p","p.signif")%>%
        add_xy_position(x = "Dosage", scales = "free")%>%
        filter(p < 0.05)
      
      subjN = length(unique(data_fortest$Subjects))
      
      # summarize mean and sd
      data_fortest_sta <- data_fortest %>%
        filter(variables == plot_variables[ivar]) %>%
        group_by(Dosage) %>%
        summarise(
          count = sum(!is.na(value)),
          sd = sd(value, na.rm = TRUE),
          se = sd / sqrt(count),
          avg = mean(value, na.rm = TRUE))
      
      
      
      
      ylim_max <- max(data_fortest_sta$avg) + 1.1 * max(data_fortest_sta$se)
  ylim_min <- min(data_fortest_sta$avg) - 1.1 * max(data_fortest_sta$se)
      
  colors_network_plot <- c(colors_network[ivar], colors_network[ivar], colors_network[ivar])
  
     myplot <- data_fortest_sta %>%
        ggplot(aes(x = Dosage, y = avg)) +
        geom_col(aes(group = Dosage, fill = Dosage), alpha = 1, size = 0.25) +
        geom_errorbar(aes(group = Dosage, color = Dosage, ymin = avg-se, ymax = avg+se), width = 0, size = 0.25) + guides(color = guide_legend(override.aes= list(alpha = 0.5))) + scale_x_discrete(breaks = waiver(), labels = xlabel_text) + scale_y_continuous(name="Size of effect", limits=c(-0.3, 0.75), breaks = c(-0.25, 0, 0.25, 0.5, 0.75), expand = c(0, 0)) + theme_classic() + theme(axis.ticks = element_line(size = 0.2), axis.line = element_line(size = 0.2),
    axis.text.x = element_text(size = 9), 
        axis.title.y=element_blank(),
        strip.text.x = element_text(size = 8, colour = "black"),
        panel.margin = unit(0.25, "lines"), legend.position="bottom", legend.margin=margin(t=-10),
        legend.spacing.y = unit(-0.05, 'cm')) + scale_colour_manual(values = colors_network_plot) + scale_fill_manual(values = colors_network_plot) + guides(color = guide_legend(override.aes= list(alpha = 0.5))) + guides(col = guide_legend(nrow = 1))  + ggtitle(plot_variables[ivar])
      if (ylim_min < 0) {
      myplot <- myplot + geom_hline(yintercept = 0, size = 0.1)
      }
      
      xtitle_size = 6
      title_size = 6
      legend_title_size = 3
      myplot <- add_theme(myplot, xtitle_size, title_size, legend_title_size)
      myplot <- myplot+scale_colour_manual(values = colors_network_plot) + scale_fill_manual(values = colors_network_plot)
      leg <- get_legend(myplot)
      plot_leg <- as_ggplot(leg)
      myplot <- myplot + theme(legend.position = "off")
      print(myplot)
    }

    
  }
  
}

```

